---
title: "'Gender Hierarchy in Fiction: A Replication of Kraicer and Piper, 2019: Social Characters: The Hierarchy of Gender in Contemporary English-Language Fiction'"
author: "Hannah Hardenbergh"
date: "5/10/2019"
output: 
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE)

# Originally, the authors organized their code with setwd() in each code chunk
# so that they didn't have to manually code all the file paths. Turns out that
# was more of a nuisance rather than a useful method. you win some, you lose
# some.

# As the author's readme says, you will need the functions and objects listed in
# this sourced file:

source("./books.data/hierarchy_adj.r")

# The author's readme ("README.txt") has great information on different
# dataframes as well as how the replication code works. This was a plus when
# working with this code.

```

This file contains code that replicates the first table and first four figures of Kraicer and Piper's report on how gender bias has found its way into contemporary fiction. Published in the Journal of Cultural Analytics on January 30, 2019, Kraicer and Piper's research is at the intersection of both the humanities and a quantitative analysis of the social sciences. Their work is extensive and impressive. The scope of this replication is to enhance the first four figures and first two tables, which will give the paper as a whole a more understandable visual for its readers. Kraicer and Piper establish several statistically significant assertions in this paper; my goal is to simply provide readers a more accurate visual for the contextualization of the underrepresentation of female-gendered characters in contemporary novels so that Kraicer and Piper's foundation for their statistical analyses is crystal-clear. Their paper is strong, and their dataset is massive. It could benefit from continued cleaning and simplifying. This is certainly an area for future research.

```{r}

# load libraries

library(epiR)
library(stats)
library(igraph)

library(dplyr)
library(bazar)
library(reshape2)
library(stringr)
library(splitstackshape)
library(rapport)

library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(gt)
library(knitr)
library(kableExtra)

library(tidyverse)

```

```{r data prep}

# unzip the folders nodes_ALL and edges_ALL.

unzip("./books.data/edges_ALL.zip")
unzip("./books.data/nodes_ALL.zip")

# get rid of all working directory calls. manually set path directories for each
# file read-in. when you get rid of it, just be careful about where you are
# trying to grab files from. Sometimes they are in different places--this
# replication code should absolve anyone trying to run this code of that
# problem.

# This function is used to turn factor variables into numeric. You can't do that
# by just going straight to numeric because of the way that factors are
# represented internally in R, though. num_func is used throughout the code.

num_func  <- function(x){as.numeric(as.character(x))}

# this function is used for figure 2- running a Rogan and Gladen's approach to
# testing for the true prevalence of how many women are included in the dataset.

epi_add  <- function(pos, tested, se, sp){
  epi  <- (epi.prev(pos, tested, se_2, sp_2)$tp[[1]])*.01
  add  <- num_func(round((tested*epi)-pos))
  return(add)
}

# metadata. just double check to make sure the file path name is correct in your
# r project.

# Every novel was also given metadata. This includes manually labelled for author
# gender as M (man), W (woman) or X (non-binary, either because of gender
# identity or multiple authorial signatures). The novels were then assigned a
# point of view (POV) of either 1p (first person) or 3p (third person) based on a
# machine learning algorithm with SVM classification based on pronoun
# distributions that was developed for this project using Python and R. The
# results of this are reported in the paper, and the code can be made available
# by contacting Eve Kraicer.

# the variable "ag", author gender, also has "T". what does this mean?


metadata.df <- data.frame(read_csv('books.data/hierarchy_metadata.csv'))

# subset by works with main character identified and by women

meta_mc.df <- metadata.df[metadata.df$mc_calc == 'Y',]
meta_w.df  <- metadata.df[metadata.df$ag == 'F',]

# tokens, nodes and edges

# interesting that I don't have the tokens file in the replication code/data.
# What/Where is it? And why can I run things without it? Is it necessary?
# Questions for the authors... for now, I commented out this line. Further,
# nowhere else in the original code is "nodes_list" or "edges_list" used. These
# objects seem like a way to organize genres of the novels that were never used.

# tokens_files <- list.files('tokens_ALL')

nodes_list <- c('nodes_YA', 'nodes_NYT', 'nodes_MY', 'nodes_SF',
                'nodes_BS', 'nodes_ROM', 'nodes_PW')

edges_list <- c('edges_YA', 'edges_NYT', 'edges_MY', 'edges_SF', 
                'edges_BS', 'edges_ROM', 'edges_PW')
```

Below I replicate the table that describes the dataset that Kraicer and Piper are working with. They selected novels from seven different popular novels lists, based on selected books for the Readers' Choice Awards, books that received prominent literary awards,  books that appeared most frequently in the New York Times's bestsellers list since 2000, and novels that were reviewed by the NYT between 2010 and 2015. Note the differences in Percentage of women authors for each category.

Interestingly, even though the authors report the effects that genre and author gender can have on gender bias in contemporary novels, the authors do not include the Romance category in their discussion of author gender effects - the data has 98% female authors for this category. Their paper mentions this once, but does not analyze the reasons for this or the implications it may have on Romance novel readers.

```{r table 1}

# gather info for table 1

# originally, authors used nlevels() and levels() to access the number of genres
# there are to build dataframes. but you need a factor to do that. the variable
# metadata.df$genre is a character string. so, I keep this for loop to build
# data frames but I used a new object, called "genre_levels".

genre_levels <- unique(metadata.df$genre)

data_info <- NULL

# list the data from the correct file path/working directory:

files <- list.files('books.data/nodes_ALL')

for (i in 1:length(genre_levels)){
  
  meta_sub <- metadata.df[metadata.df$genre == genre_levels[i],]
  
  # Start to make columns with data
  
  titles  <- files[files %in% meta_sub$ID]
  author <- str_split_fixed(titles, "_", 3)
  no_authors <- length(unique(author[,2]))
  genre <- genre_levels[i]
  novels <- length(titles)
  perc_w <- (length(which(titles %in% meta_w.df$ID))/novels)*100
  
  # bind for a dataframe
  
  temp_info <- data.frame(cbind(genre, novels, no_authors, perc_w))
  data_info <- data.frame(rbind(data_info, temp_info))
}

# build and order table 1 

table_1 <- data_info[order(num_func(data_info$perc_w)),]

# make it pretty: column names, round percentages, etc. using gt().

table_1_tidy <- table_1 %>% 
  rename("code" = "genre") %>% 
  mutate(genre = c("Science Fiction",
                   "Prizewinners",
                   "Bestsellers",
                   "NYT Reviewed",
                   "Mystery",
                   "Young Adult",
                   "Romance"))

table_1_tidy %>%
  mutate(perc_z = as.numeric(as.character(perc_w))) %>% 
  select(code, novels, no_authors, perc_z, genre) %>% 
  gt() %>% 
  cols_move_to_start(columns = vars(genre)) %>% 
  cols_label(code = "Code", 
             novels = "Novels",
             no_authors = "# Authors",
             perc_z = "% Women Authors",
             genre = "Genre") %>% 
  tab_footnote("Table 1. Summary of our data, with the number of works per genre, number of authors. Genres are ranked by percent of women authors per genre.",
               locations = cells_column_labels(columns = vars(genre))) %>% 
  fmt_number(columns = vars(perc_z), decimals = 2)


```
------

The authors use the first set of figures to contextualize their research and set up their dataset for later analysis. They simply explain how their character rankings method correlates to the number of times a character is mentioned within 100,000 words of a selected text in any of their novels in the first two figures. The figures are replicated below, however it would be much simpler for whoever is reading the original paper just to get a clearer visual of something more interesting other than a black line that simply shows the nature of a ranked list. The correlation between a ranked list and the number of times a character is mentioned in a novel would be easy to understand with a short explanation included in the paper. Therefore, if this paper were to be revised, I would include simple bar graphs to show the imbalance of gender in characters who are mentioned more often within text. It becomes clear that 1) male characters are more prominent in contemporary text, and 2) several characters in the authors' dataset are not assigned male or female. *Visualizing* these details would be a great addition to contextualizing the type of gender bias that the authors begin to work with in their later tests to find more detailed conclusions about the effect of gender in contemporary fiction.





```{r}
# 1c: FIGURE ONE
# - - - - - - - - 

# DATA
# calculates the mentions for each character by rank position 

# run twice turning off / on condition
# once on all works (mentions) 
# once on only works with main character identified by gender (mentions_sub)

mentions <- NULL

# mentions_sub <- NULL

# set working directory to where the data lives.

files <- list.files('./books.data/nodes_ALL')

# this for loop is grabbing the number of counts that a character is mentioned
# in the BookNLP data found in the "nodes_ALL" folder.

for(j in 1:length(files)){
  work  <- files[j]
  nodes <- data.frame(read_csv(paste0('books.data/nodes_ALL/', (files[j]))))

  # only pull mentions
  
  m_counts <- data.frame(nodes[,2]) 
  
  # add columns
  
  extra <- as.data.frame(matrix(c(rep.int(NA,length(200))),nrow=200))
  
  #give same column names
  
  colnames(extra) <- colnames(m_counts) 
  
  #bind
  
  mentions_bind <- rbind(m_counts, extra) 
  mentions <- cbind(mentions, mentions_bind[1:150,])
  
  # mentions_sub <- cbind(mentions_sub, mentions_bind[1:150,])
  # }
}

# count up sums for each rank position for both cases 

rank_m <- NULL
rank_m_sub <- NULL

for (i in 1:nrow(mentions)){
  mean <- mean(num_func(mentions[i,]), na.rm = TRUE)
  sum <- sum(mentions[i,], na.rm = TRUE)
  log_sum <- log(sum)
  rank <- i
  temp <- data.frame(cbind(rank, mean, sum, log_sum))
  rank_m <- data.frame(rbind(rank_m, temp))
}

# for (i in 1:nrow(mentions_sub)){
#   mean <- mean(num_func(mentions_sub[i,]), na.rm=TRUE)
#   sum <- sum(mentions_sub[i,], na.rm=TRUE)
#   log_sum <- log(sum)
#   rank <- i
#   temp <- data.frame(cbind(rank,mean,sum,log_sum))
#   rank_m_sub <- data.frame(rbind(rank_m_sub,temp))
# }

# BUILD  

# run on all works 

fig_1 <- ggplot(data=rank_m, aes(x=rank, y=mean)) +
  geom_point()+
  ggtitle("Mean Mentions by Rank")+
  labs(x="Character Rank by Mentions", y="Mean Mentions")+
  theme_bw()+
  geom_vline(xintercept = 20, linetype = "dashed") +
  theme(plot.title = element_text(size = 13), 
                                  # family = "Times New Roman", face = "bold"),
        text = element_text(size = 11), 
                            # family = "Times New Roman"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11))

print(fig_1)

ggsave(filename = "figure_1.jpg", plot = fig_1, path = "./figures/", width = 7, height = 7)
```


Below is an extension and improvement of the above figure. 

```{r}
## Figure 1 Extension part 1

#color is always nice too!

nodes_tidy <- nodes %>% 
  rename(num_mentions = X..MENTIONS...100K.WORDS,
         name = NAME,
         gender = GENDER) %>% 
  arrange(desc(gender))

fig_1_ext <- nodes_tidy %>% 
  #filter(gender == "M") %>% 
  ggplot(aes(x = gender, y = num_mentions, fill = gender)) + geom_col() + 
  ylab("Gender") + xlab("Mentions per 100k Words")

print(fig_1_ext)

```

------

The authors describe their "manual" process of assigning gender ID's to each character once they acquired the data. This process included an error based on how their methods failed to assign some (about fifteen) characters either male or female. Below is, first, the replication of the authors' graph that shows that as character mentions increase, the number of "unassigned" gendered characters decreases. However, it would be more interesting to simply focus on how significantly the unassigned characters effect the number of times a character is mentioned. 

I then include another table and bar graphs to visualize this error. It is easier to understand that, by this bar graph, should the number of unassigned (marked with a "?" in the data) gendered characters be female, the number of female character mentions would be greater. Additionally, visualizing this error also calls more attention to the question of using a gender binary in contemporary fiction. What if there were other categories besides male and female?

```{r }
# 1d: FIGURE TWO
# - - - - - - - - 

# DATA
# calculates raw counts of M/ W/ ? by rank position 

files <- list.files('./books.data/nodes_ALL')

# run four times turning off / on condition
# once on all works (raw) 
# once on only works with main character identified by gender (raw_sub)
# once on all works by women (raw_w) 
# once on only works by women with main character identified by gender (raw_sub_w)

# the authors always make empty objects for future use. Good practice is to make
# a vector with accurate dimensions, if you know them. But the authors
# just create objects with empty space to use later, using NULL.

raw <- NULL
raw_sub <- NULL
raw_w  <- NULL
raw_sub_w <- NULL

for(i in 1:length(files)){ 
  work  <- files[i]
  
  # only files where top pair was calculated 
  
  if (work %in% meta_mc.df$ID){
    
    # only files written by women authors 
    
    if (work %in% meta_w.df$ID){
      
      nodes <- data.frame(read_csv(paste0('books.data/nodes_ALL/', (files[i]))))
      genders <- data.frame(nodes[,3]) 
      
      # add columns to cbind different books 
      
      extra <- as.data.frame(matrix(c(rep.int(NA,length(200))),nrow=200)) 
      colnames(extra) <- colnames(genders)
      genders_bind <- rbind(genders, extra) 
      
      raw <- data.frame(cbind(raw, genders_bind[1:150,]))
      raw_sub <- data.frame(cbind(raw_sub, genders_bind[1:150,]))
      raw_w <- data.frame(cbind(raw_w, genders_bind[1:150,]))
      raw_sub_w  <- data.frame(cbind(raw_sub_w, genders_bind[1:150,]))
    }
  }
}

# clean vectors to normalize factors --> characters. This function will make
# gender identifiers. It does so by using gsub(), which replaces the numbers 1,
# 2, or 3 with 'M', 'F', or '?'.

clean_func  <- function(df){
  v <- df[,1]
  clean <- gsub(3, 'M', v)
  clean <- gsub(2, 'F', clean)
  clean <- gsub(1, '?', clean)
  return(clean)
}

raw[,1] <- clean_func(raw)
raw_sub[,1] <- clean_func(raw_sub)
raw_w[,1] <- clean_func(raw_w)
raw_sub_w[,1] <- clean_func(raw_sub_w)

# count up sums for each rank position
x <- NULL

rank_func  <- function(x){
  ranks <- NULL
  for (i in 1:nrow(x)){
    wr <- length(which(x[i,] == 'F')) # how many women
    mr <- length(which(x[i,] == 'M')) # how many men
    q  <- length(which(x[i,] == '?')) # how many unknown
    
    all <- mr + wr + q
    
    # adjust women and men 
    
    women  <- wr + epi_add(wr,
                           all,
                           se_2, 
                           sp_2)
    men  <- mr + epi_add(mr,
                         all,
                         se_2m,
                         se_2m)
    temp <- data.frame(women,
                       men,
                       q,
                       all) 
    #cbind for storage
    ranks <-data.frame(rbind(ranks,
                             temp))
  }
  
  ranks$rank <- 1:150
  ranks$cumulative_q  <- cumsum(ranks$q)/cumsum(ranks$all)
  return(ranks)
}

# here, the code is using the rank_func function that the authors create in
# hierarchy_adj.r, which is sourced at the beginning of this code. It makes the
# estimates for true prevalence tests that validate that some gender assignments
# in books are inaccurate. as in, on page 8 of the paper when authors discuss
# the proportions of accurately assigned and inaccurately assigned gender
# identifiers (male, female, or unknown.) Basically, this means that if there is
# a higher sensitivity rate for one gender, that means they have been assigned
# more accurately. when men have this higher rate, it accounts for the
# possibility of women being undercounted in the dataset after they have
# assigned gender identifiers manually.

# in order to run properly, apparent prevalence can't be less than (1 - Sp),
# otherwise the estimate of true prevalance will be "invalid".

ranks_all <- rank_func(raw)
ranks_sub  <- rank_func(raw_sub)
ranks_w  <- rank_func(raw_w)                   
ranks_sub_w  <- rank_func(raw_sub_w)   

# basic counts by subsets
# how many unknown overall?

perc_unknown10 <- sum(ranks_all$q[1:10])/sum(ranks_all$all[1:10])
perc_unknown20 <- sum(ranks_all$q[1:20])/sum(ranks_all$all[1:20])

# BUILD 
# run on all works

figure_2 <- ggplot(data=ranks_all, aes(x = rank, y = cumulative_q)) +
  geom_line()+
  ggtitle("% Unassigned Characters by Mentions Rank")+
  labs(x = "Character Rank by Mentions", 
       y = "Cumulative Percent Unassigned")+
  theme_bw() + 
  geom_vline(xintercept = 20, linetype = "dashed") + 
  theme(plot.title = element_text(size = 13, 
                                  # family = "Times New Roman", 
                                  face = "bold"),
        text = element_text(size = 11), 
                            #family = "Times New Roman"),
        axis.title = element_text(face="bold"),
        axis.text.x = element_text(size = 11))

print(figure_2)

ggsave(filename = "figure_2.jpg", plot = figure_2, path = "./figures/", width = 7, height = 7)

```

```{r, results='asis'}

# show quickly the numbers for each character ID:

gender_count <- nodes_tidy %>% 
  select(gender) %>% 
  count(gender) %>% 
  gt() %>% 
  tab_header(title = "Number of Gender Assignments for Male, Female, and Unknown") %>% 
  cols_label(gender = "Gender",
             n = "Number of Characters")


# now make a figure that simply shows the difference between number of mentions
# between genders of characters.

fig_1_ext_mf <- nodes_tidy %>% 
  filter(gender != "?") %>% 
  ggplot(aes(x = gender, y = num_mentions, fill = gender)) + geom_col() + 
  ylab("Character Mentions (Per 100k words)") + 
  xlab("Gender of Character") + 
  ggtitle("Discrepancies of Gender in Character Mentions")

print(fig_1_ext_mf)
```

```{r }

# extension part 2

# make a figure that will show that the unidentified characters constitute for
# about another 50% of how many times female characters are mentioned, which is
# the error that the authors talk about in their paper. if these were identified
# as female, it would obviously change the way we see the number of times a
# gendered character is mentioned in 100K words. see my paper for a further
# discussion of how creating this binary may actually be less beneficial for
# future studies, and considering the challenge of moving past this binary in
# this quantitative analysis.

fig_1_ext_f <- nodes_tidy %>% 
  filter(gender != "M") %>% 
  ggplot(aes(x = gender, y = num_mentions, fill = gender)) + geom_col() + 
  ylab("Gender of Character") +
  xlab("Character Mentions (Per 100k words)")

print(fig_1_ext_f)

```



```{r  include = FALSE}
# ------------------------
# SECTION II: VISIBILITY
# ------------------------


# bootstrapper 
# x is column on df to be sampled, y is df

boot_func <- function(x,y){
  sample <- as.character(sample(x, nrow(y), replace = TRUE))
  return(sample)
}

# adjustments
# for all
epi_func <- function(vector,w,m,q,se,sp){
  pos <- length(which(as.character(vector) == w))
  tested <- length(which(as.character(vector) == w | as.character(vector) == m | as.character(vector) == q)) 
  epi <- epi.prev(pos, tested, se, sp)
  adj_perc <- epi$tp[[1]]*.01
  adj_add <- round((tested*adj_perc)-pos)
  return(adj_add)
}

# for main chars
epi_func2 <- function(vector,w,m,se,sp){
  pos <- length(which(as.character(vector) == w))
  tested <- length(which(as.character(vector) == w | as.character(vector) == m)) # original sample set 
  epi <- epi.prev(pos, tested, se, sp)
  # find percentage increase of women for true prevalence of women as main character
  adj_perc <- epi$tp[[1]]*.01
  # how many more women is that? 
  adj_add <- round((tested*adj_perc)-pos)
  return(adj_add)
}

# 1: TOP 20 & ALL 
# = = = = = = = = = 

# build master table with genders of top pairs from each document

# run twice for conditionals 
# once for all 
# once for only characters in the top twnety 

chars_all  <- NULL
chars_20 <- NULL

for (i in 1:length(genre_levels)){
  
  #loop through each file
  meta_sub <- metadata.df[metadata.df$genre == genre_levels[i],]
  titles  <- files[files %in% meta_sub$ID]
  
  # loop through each file
  chars <- NULL
  for(j in 1:length(titles)){ 
    work <- titles[j]
    nodes <- data.frame(read_csv(paste0('./books.data/nodes_ALL/', (titles[j]))))

    
    # genders  <- as.character(nodes[,3])
    # for 20 chars run these lines 
    # make sure work has 20 chars
    
    if (nrow(nodes) >= 20){
      genders <- as.character(nodes[1:20,3])
    } else {
      genders <- as.character(nodes[,3])}
    ag <- as.character(metadata.df$ag[(which(metadata.df$ID %in% work))])
    genre <- genre_levels[i]
    
    temp_chars <- data.frame(cbind(work, genders, ag, genre))
    chars <- data.frame(rbind(temp_chars, chars))
  }
  chars_20  <- data.frame(rbind(chars_20, chars))
  
  # chars_all <- data.frame(rbind(chars_all, chars))
}
```


```{r  include = FALSE}
# 1a: ADJUSTMENTS
# - - - - - - - - - 

# how many M / ? need to be replaced with Fs?

men_rep2  <- epi_func(chars_20$genders, 'M', 'F', '?', se_2m, sp_2m)
q_rep2  <- (epi_func(chars_20$genders, 'F', 'M', '?', se_2, sp_2))-abs(men_rep2)

# randomly replace M 

rep_m20 <- sample(which(chars_20$gender == 'M'), abs(men_rep2))
chars_20$adj_all <- replace(as.character(chars_20$gender), rep_m20, 'F')

# randomly replace ?

rep_q20 <- sample(which(chars_20$adj_all == '?'), q_rep2)
chars_20$adj_all <- replace(as.character(chars_20$adj_all), rep_q20, 'F')


# BOOTSTRAP 
# calculate the percentage of 'F' with bootstrapping 

boot_chars_20 <- vector()
boot_chars_all  <- vector()
for (j in 1:10000){  
  sample <- boot_func(chars_20$adj_all, chars_20)
  # sample  <- boot_func(chars_all$adj_all, chars_all)
  f_perc <- length(which(sample == 'F'))/length(sample)
  boot_chars_20[j] <- f_perc
  # boot_chars_all[j] <- f_perc
}

```

```{r  include = FALSE}
# 1b: REPORTED VALUES
# - - - - - - - - - - 

length(which(chars_20$genders == 'F')) # raw women
length(which(chars_20$adj_all == 'F')) # adjusted women 
length(which(chars_20$adj_all == 'F')) / nrow(chars_20) # % women 
mean(boot_chars_20) # bootstrap mean - top 20 
mean(boot_chars_all) # bootstrap mean - all 

# by genre and ag 

chars_func  <- function(x){
  results <- NULL
  for (i in 1:nlevels(x)){
    sub <- chars_20[which(as.character(x) == levels(x)[i]),]
    genre <- levels(x)[i]
    f <- vector()
    for (j in 1:10000){  
      sample <- boot_func(sub$adj_all, sub)
      f[j] <- length(which(sample == 'F'))
    }
    all  <- nrow(sub)
    temp <- data.frame(cbind('f'= mean(f), 'all'= all, genre))
    results <- data.frame(rbind(results,temp))
  }
  results$f_perc  <- num_func(results$f)/num_func(results$all)
  return(results)
}

# by genre and author gender 

boot_chars_genre <- chars_func(chars_20$genre)

# remove romance

chars_20xROM <- chars_20[which(chars_20$genre != 'ROM'),]
boot_chars_ag <- chars_func(chars_20xROM$ag)
boot_chars_ag_wROM <- chars_func(chars_20$ag)
```


```{r include=FALSE}
# 1d: SIGNIFICANCE TESTING
# - - - - - - - - - - - - - 

chisq_20  <- t(data.frame(cbind('women' = num_func(boot_chars_genre$f), 
                                'all'=num_func(boot_chars_genre$all))))
chisq.test(chisq_20)

chisq_20_ag  <- t(data.frame(cbind('women' = num_func(boot_chars_ag$f[1:2]),
                                   'all'=num_func(boot_chars_ag$all[1:2]))))
chisq.test(chisq_20_ag)
fisher.test(chisq_20_ag)
```


```{r results='asis'}

# TABLE TWO

table_2  <- data.frame(cbind('genre' = as.character(boot_chars_genre$genre),
                             'percent_w' = num_func(boot_chars_genre$f_perc)*100))

table_2 <- table_2[order(num_func(table_2$percent_w)),]

table_2_tidy <- table_2 %>%
  gt() %>% 
  tab_header("Estimates for Prevalence of Women in Top-Twenty Characters in Novels of All Genres") %>% 
  cols_label(genre = "Genre",
             percent_w = "% Women Top-Twenty Characters")

ggsave(filename = "table_2.jpg", plot = tableGrob(table_2_tidy), path = "./figures/", width = 7, height = 7)

```


The following graphics show how gender in authors effects the prominence of male and female characters across genres. I replicate both of the author's plots as they were originally published. Interestingly, as the authors find, protagonist characters have about a 50:50 ratio of male to female characters, while all other main characters remain at about a 62:38 ratio of male to female. This is displayed in the plots. Additionally, based on their statistical methods and tests, Kraicer and Piper find that if the author is a woman, they are more likely to include a woman protagonist. The data points to the left of the dotted line show the top-20 ranked characters (i.e. most mentioned) in each novel.

```{r results='asis'}

# figure 3

# from ranks_sub, update the first and second character based on perspective calcs

ranks_sub[1,1:2]  <- c(623, 617)
ranks_sub[2,1:2]  <- c(470, 768)
ranks_sub_w[1,1:2]  <- c(321,194)
ranks_sub_w[2,1:2]  <- c(191,324)

# get correct percentages of women 

ranks_sub$perc_w  <- ranks_sub$w/ranks_sub$all
ranks_sub_w$perc_w  <- ranks_sub_w$w/ranks_sub_w$all

# BUILD 
# figure 3 
# uses only works with main character identified 

figure_3 <- ggplot(data=ranks_sub, aes(rank)) + 
  geom_line(aes(y = women, colour = "Women")) + 
  geom_line(aes(y = men, colour = "Men")) +
  scale_color_manual(values=c("darkorange2", "mediumpurple4")) +
  ggtitle("Men and Women by Rank Position (All Authors)")+
  labs(x="Rank of Character", y="") +
  labs(color="Character Gender") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, 
                                  # family = "Times New Roman", 
                                  face = "bold"),
        text = element_text(size = 11), 
                            # family = "Times New Roman"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) +
  geom_vline(xintercept = 20, linetype = "dashed") 

print(figure_3)

ggsave(filename = "figure_3.jpg", plot = figure_3, path = "./figures/", width = 7, height = 7)

```


``` {r results='asis'}

# figure 4 

figure_4 <- ggplot(data = ranks_sub_w, aes(rank)) + 
  geom_line(aes(y = women, colour = "Women")) + 
  geom_line(aes(y = men, colour = "Men")) +
  scale_color_manual(values=c("darkorange2", "mediumpurple4")) +
  ggtitle("Men and Women by Rank Position (Women Authors)")+
  labs(x="Rank of Character", y="") +
  labs(color="Character Gender") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, 
                                  # family = "Times New Roman", 
                                  face = "bold"),
        text = element_text(size = 11), 
                            # family = "Times New Roman"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) +
  geom_vline(xintercept = 20, linetype = "dashed") 

print(figure_4)

ggsave(filename = "figure_4.jpg", plot = figure_4, path = "./figures/", width = 7, height = 7)

```


References

REFERENCES
1.	David Bamman, Ted Underwood and Noah Smith, "A Bayesian Mixed Effects Model of Literary Character," ACL 2014.
2.	Eve Kraicer and Andrew Piper, “Social Characters: The Hierarchy of Gender in Contemporary English-Language Fiction,” Journal of Cultural Analytics. January 30, 2019.
