---
title: "Gender Hierarchy in Fiction: A Replication of Kraicer and Piper, 2019: Social Characters: The Hierarchy of Gender in Contemporary English-Language Fiction"
author: "Hannah Hardenbergh"
date: "5/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      error = FALSE)

# Originally, the authors organized their code with setwd() in each code chunk
# so that they didn't have to manually code all the file paths. Turns out that
# was more of a nuisance rather than a useful method. you win some, you lose
# some.

# As the author's readme says, you will need the functions and objects listed in
# this sourced file:

source("./books.data/hierarchy_adj.r")

# The author's readme ("README.txt") has great information on different
# dataframes as well as how the replication code works. This was a plus when
# working with this code.

```

This file contains code that replicates the first table and first four figures of Kraicer and Piper's report on how gender bias has found its way into contemporary fiction. Published in the Journal of Cultural Analytics on January 30, 2019, Kraicer and Piper's research is at the intersection of both the humanities and a quantitative analysis of the social sciences. Their work is extensive and impressive. The scope of this replication is to enhance the first four figures and first two tables, which will give the paper as a whole a more understandable visual for its readers. Kraicer and Piper establish several statistically significant assertions in this paper; my goal is to simply provide readers a more accurate visual for the contextualization of the underrepresentation of female-gendered characters in contemporary novels so that Kraicer and Piper's foundation for their statistical analyses is crystal-clear. Their paper is strong, and their dataset is massive. It could benefit from continued cleaning and simplifying. This is certainly an area for future research.

```{r}

# load libraries

library(epiR)
library(stats)
library(igraph)

library(dplyr)
library(bazar)
library(reshape2)
library(stringr)
library(splitstackshape)
library(rapport)

library(ggplot2)
library(gridExtra)
library(RColorBrewer)
library(gt)
library(knitr)
library(kableExtra)

library(tidyverse)

```

```{r data prep}

# unzip the folders nodes_ALL and edges_ALL.

unzip("./books.data/edges_ALL.zip")
unzip("./books.data/nodes_ALL.zip")

# get rid of all working directory calls. manually set path directories for each
# file read-in. when you get rid of it, just be careful about where you are
# trying to grab files from. Sometimes they are in different places--this
# replication code should absolve anyone trying to run this code of that
# problem.

# This function is used to turn factor variables into numeric. You can't do that
# by just going straight to numeric because of the way that factors are
# represented internally in R, though. num_func is used throughout the code.

num_func  <- function(x){as.numeric(as.character(x))}

# this function is used for figure 2- running a Rogan and Gladen's approach to
# testing for the true prevalence of how many women are included in the dataset.

epi_add  <- function(pos, tested, se, sp){
  epi  <- (epi.prev(pos, tested, se_2, sp_2)$tp[[1]])*.01
  add  <- num_func(round((tested*epi)-pos))
  return(add)
}

# metadata. just double check to make sure the file path name is correct in your
# r project.

# Every novel was also given metadata. This includes manually labelled for author
# gender as M (man), W (woman) or X (non-binary, either because of gender
# identity or multiple authorial signatures). The novels were then assigned a
# point of view (POV) of either 1p (first person) or 3p (third person) based on a
# machine learning algorithm with SVM classification based on pronoun
# distributions that was developed for this project using Python and R. The
# results of this are reported in the paper, and the code can be made available
# by contacting Eve Kraicer.

# the variable "ag", author gender, also has "T". what does this mean?


metadata.df <- data.frame(read_csv('books.data/hierarchy_metadata.csv'))

# subset by works with main character identified and by women

meta_mc.df <- metadata.df[metadata.df$mc_calc == 'Y',]
meta_w.df  <- metadata.df[metadata.df$ag == 'F',]

# tokens, nodes and edges

# interesting that I don't have the tokens file in the replication code/data.
# What/Where is it? And why can I run things without it? Is it necessary?
# Questions for the authors... for now, I commented out this line. Further,
# nowhere else in the original code is "nodes_list" or "edges_list" used. These
# objects seem like a way to organize genres of the novels that were never used.

# tokens_files <- list.files('tokens_ALL')

nodes_list <- c('nodes_YA', 'nodes_NYT', 'nodes_MY', 'nodes_SF',
                'nodes_BS', 'nodes_ROM', 'nodes_PW')

edges_list <- c('edges_YA', 'edges_NYT', 'edges_MY', 'edges_SF', 
                'edges_BS', 'edges_ROM', 'edges_PW')
```

```{r table 1}

# gather info for table 1

# originally, authors used nlevels() and levels() to access the number of genres
# there are to build dataframes. but you need a factor to do that. the variable
# metadata.df$genre is a character string. so, I keep this for loop to build
# data frames but I used a new object, called "genre_levels".

genre_levels <- unique(metadata.df$genre)

data_info <- NULL

# list the data from the correct file path/working directory:

files <- list.files('books.data/nodes_ALL')

for (i in 1:length(genre_levels)){
  
  meta_sub <- metadata.df[metadata.df$genre == genre_levels[i],]
  
  # Start to make columns with data
  
  titles  <- files[files %in% meta_sub$ID]
  author <- str_split_fixed(titles, "_", 3)
  no_authors <- length(unique(author[,2]))
  genre <- genre_levels[i]
  novels <- length(titles)
  perc_w <- (length(which(titles %in% meta_w.df$ID))/novels)*100
  
  # bind for a dataframe
  
  temp_info <- data.frame(cbind(genre, novels, no_authors, perc_w))
  data_info <- data.frame(rbind(data_info, temp_info))
}

# build and order table 1 

table_1 <- data_info[order(num_func(data_info$perc_w)),]

# make it pretty: column names, round percentages, etc. using gt().

table_1_tidy <- table_1 %>% 
  rename("code" = "genre") %>% 
  mutate(genre = c("Science Fiction",
                   "Prizewinners",
                   "Bestsellers",
                   "NYT Reviewed",
                   "Mystery",
                   "Young Adult",
                   "Romance"))

table_1_tidy %>%
  mutate(perc_z = as.numeric(as.character(perc_w))) %>% 
  select(code, novels, no_authors, perc_z, genre) %>% 
  gt() %>% 
  cols_move_to_start(columns = vars(genre)) %>% 
  cols_label(code = "Code", 
             novels = "Novels",
             no_authors = "# Authors",
             perc_z = "% Women Authors",
             genre = "Genre") %>% 
  tab_footnote("Table 1. Summary of our data, with the number of works per genre, number of authors. Genres are ranked by percent of women authors per genre.",
               locations = cells_column_labels(columns = vars(genre))) %>% 
  fmt_number(columns = vars(perc_z), decimals = 2)

```


```{r }

# 1c: FIGURE ONE
# - - - - - - - - 

# DATA
# calculates the mentions for each character by rank position 

# run twice turning off / on condition
# once on all works (mentions) 
# once on only works with main character identified by gender (mentions_sub)

mentions <- NULL

# mentions_sub <- NULL

# set working directory to where the data lives.

files <- list.files('./books.data/nodes_ALL')

for(j in 1:length(files)){
  work  <- files[j]
  nodes <- data.frame(read_csv(paste0('books.data/nodes_ALL/', (files[j]))))

  # only pull mentions
  
  m_counts <- data.frame(nodes[,2]) 
  
  # add columns
  
  extra <- as.data.frame(matrix(c(rep.int(NA,length(200))),nrow=200))
  
  #give same column names
  
  colnames(extra) <- colnames(m_counts) 
  
  #bind
  
  mentions_bind <- rbind(m_counts, extra) 
  mentions <- cbind(mentions, mentions_bind[1:150,])
  
  # mentions_sub <- cbind(mentions_sub, mentions_bind[1:150,])
  # }
}

# count up sums for each rank position for both cases 
rank_m <- NULL
rank_m_sub <- NULL

for (i in 1:nrow(mentions)){
  mean <- mean(num_func(mentions[i,]), na.rm=TRUE)
  sum <- sum(mentions[i,], na.rm=TRUE)
  log_sum <- log(sum)
  rank <- i
  temp <- data.frame(cbind(rank,mean,sum,log_sum))
  rank_m <- data.frame(rbind(rank_m,temp))
}

# for (i in 1:nrow(mentions_sub)){
#   mean <- mean(num_func(mentions_sub[i,]), na.rm=TRUE)
#   sum <- sum(mentions_sub[i,], na.rm=TRUE)
#   log_sum <- log(sum)
#   rank <- i
#   temp <- data.frame(cbind(rank,mean,sum,log_sum))
#   rank_m_sub <- data.frame(rbind(rank_m_sub,temp))
# }

# BUILD  

# run on all works 

fig_1 <- ggplot(data=rank_m, aes(x=rank, y=mean)) +
  geom_point()+
  ggtitle("Mean Mentions by Rank")+
  labs(x="Character Rank by Mentions", y="Mean Mentions")+
  theme_bw()+
  geom_vline(xintercept = 20, linetype = "dashed") +
  theme(plot.title = element_text(size = 13), 
                                  # family = "Times New Roman", face = "bold"),
        text = element_text(size = 11), 
                            # family = "Times New Roman"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11))

ggsave(filename = "figure_1.jpg", plot = fig_1, path = "./figures/", width = 7, height = 7)
```


```{r }

# 1d: FIGURE TWO
# - - - - - - - - 

# DATA
# calculates raw counts of M/ W/ ? by rank position 

files <- list.files('./books.data/nodes_ALL')

# run four times turning off / on condition
# once on all works (raw) 
# once on only works with main character identified by gender (raw_sub)
# once on all works by women (raw_w) 
# once on only works by women with main character identified by gender (raw_sub_w)

# what you should do is make a vector with accurate dimensions if you know it,
# with NA's. but the authors just create objects with empty space to use later
# using NULL. 

raw <- NULL
raw_sub <- NULL
raw_w  <- NULL
raw_sub_w <- NULL

for(i in 1:length(files)){ 
  work  <- files[i]
  
  # only files where top pair was calculated 
  
  if (work %in% meta_mc.df$ID){
    
    # only files written by women authors 
    
    if (work %in% meta_w.df$ID){
      
      nodes <- data.frame(read_csv(paste0('books.data/nodes_ALL/', (files[i]))))
      genders <- data.frame(nodes[,3]) 
      
      # add columns to cbind different books 
      
      extra <- as.data.frame(matrix(c(rep.int(NA,length(200))),nrow=200)) 
      colnames(extra) <- colnames(genders)
      genders_bind <- rbind(genders, extra) 
      
      raw <- data.frame(cbind(raw, genders_bind[1:150,]))
      raw_sub <- data.frame(cbind(raw_sub, genders_bind[1:150,]))
      raw_w <- data.frame(cbind(raw_w, genders_bind[1:150,]))
      raw_sub_w  <- data.frame(cbind(raw_sub_w, genders_bind[1:150,]))
    }
  }
}

# clean vectors to normalize factors --> characters. This function will make
# gender identifiers. It does so by using gsub(), which replaces the numbers 1,
# 2, or 3 with 'M', 'F', or '?'.

clean_func  <- function(df){
  v <- df[,1]
  clean <- gsub(3, 'M', v)
  clean <- gsub(2, 'F', clean)
  clean <- gsub(1, '?', clean)
  return(clean)
}

raw[,1] <- clean_func(raw)
raw_sub[,1] <- clean_func(raw_sub)
raw_w[,1] <- clean_func(raw_w)
raw_sub_w[,1] <- clean_func(raw_sub_w)

# count up sums for each rank position
x <- NULL

rank_func  <- function(x){
  ranks <- NULL
  for (i in 1:nrow(x)){
    wr <- length(which(x[i,] == 'F')) # how many women
    mr <- length(which(x[i,] == 'M')) # how many men
    q  <- length(which(x[i,] == '?')) # how many unknown
    
    all <- mr + wr + q
    
    # adjust women and men 
    
    women  <- wr + epi_add(wr,
                           all,
                           se_2, 
                           sp_2)
    men  <- mr + epi_add(mr,
                         all,
                         se_2m,
                         se_2m)
    temp <- data.frame(women,
                       men,
                       q,
                       all) 
    #cbind for storage
    ranks <-data.frame(rbind(ranks,
                             temp))
  }
  
  ranks$rank <- 1:150
  ranks$cumulative_q  <- cumsum(ranks$q)/cumsum(ranks$all)
  return(ranks)
}

# here, the code is using the rank_func function that the authors create in
# hierarchy_adj.r, which is sourced at the beginning of this code. It makes the
# estimates for true prevalence tests that validate that some gender assignments
# in books are inaccurate. as in, on page 8 of the paper when authors discuss
# the proportions of accurately assigned and inaccurately assigned gender
# identifiers (male, female, or unknown.) Basically, this means that if there is
# a higher sensitivity rate for one gender, that means they have been assigned
# more accurately. when men have this higher rate, it accounts for the
# possibility of women being undercounted in the dataset after they have
# assigned gender identifiers manually.

# in order to run properly, apparent prevalence can't be less than (1 - Sp),
# otherwise the estimate of true prevalance will be "invalid".

ranks_all <- rank_func(raw)
ranks_sub  <- rank_func(raw_sub)
ranks_w  <- rank_func(raw_w)                   
ranks_sub_w  <- rank_func(raw_sub_w)   

# basic counts by subsets
# how many unknown overall?

perc_unknown10 <- sum(ranks_all$q[1:10])/sum(ranks_all$all[1:10])
perc_unknown20 <- sum(ranks_all$q[1:20])/sum(ranks_all$all[1:20])

# BUILD 
# run on all works

figure_2 <- ggplot(data=ranks_all, aes(x = rank, y = cumulative_q)) +
  geom_line()+
  ggtitle("% Unassigned Characters by Mentions Rank")+
  labs(x = "Character Rank by Mentions", 
       y = "Cumulative Percent Unassigned")+
  theme_bw() + 
  geom_vline(xintercept = 20, linetype = "dashed") + 
  theme(plot.title = element_text(size = 13, 
                                  # family = "Times New Roman", 
                                  face = "bold"),
        text = element_text(size = 11), 
                            #family = "Times New Roman"),
        axis.title = element_text(face="bold"),
        axis.text.x = element_text(size = 11))

ggsave(filename = "figure_2.jpg", plot = figure_2, path = "./figures/", width = 7, height = 7)

```


```{r  include = FALSE}
# ------------------------
# SECTION II: VISIBILITY
# ------------------------

# bootstrapper 
# x is column on df to be sampled, y is df

boot_func <- function(x,y){
  sample <- as.character(sample(x, nrow(y), replace = TRUE))
  return(sample)
}

# adjustments
# for all
epi_func <- function(vector,w,m,q,se,sp){
  pos <- length(which(as.character(vector) == w))
  tested <- length(which(as.character(vector) == w | as.character(vector) == m | as.character(vector) == q)) 
  epi <- epi.prev(pos, tested, se, sp)
  adj_perc <- epi$tp[[1]]*.01
  adj_add <- round((tested*adj_perc)-pos)
  return(adj_add)
}

# for main chars
epi_func2 <- function(vector,w,m,se,sp){
  pos <- length(which(as.character(vector) == w))
  tested <- length(which(as.character(vector) == w | as.character(vector) == m)) # original sample set 
  epi <- epi.prev(pos, tested, se, sp)
  # find percentage increase of women for true prevalence of women as main character
  adj_perc <- epi$tp[[1]]*.01
  # how many more women is that? 
  adj_add <- round((tested*adj_perc)-pos)
  return(adj_add)
}

# 1: TOP 20 & ALL 
# = = = = = = = = = 

# build master table with genders of top pairs from each document

# run twice for conditionals 
# once for all 
# once for only characters in the top twnety 

chars_all  <- NULL
chars_20 <- NULL

for (i in 1:length(genre_levels)){
  
  #loop through each file
  meta_sub <- metadata.df[metadata.df$genre == genre_levels[i],]
  titles  <- files[files %in% meta_sub$ID]
  
  # loop through each file
  chars <- NULL
  for(j in 1:length(titles)){ 
    work <- titles[j]
    nodes <- data.frame(read_csv(paste0('./books.data/nodes_ALL/', (titles[j]))))

    
    # genders  <- as.character(nodes[,3])
    # for 20 chars run these lines 
    # make sure work has 20 chars
    
    if (nrow(nodes) >= 20){
      genders <- as.character(nodes[1:20,3])
    } else {
      genders <- as.character(nodes[,3])}
    ag <- as.character(metadata.df$ag[(which(metadata.df$ID %in% work))])
    genre <- genre_levels[i]
    
    temp_chars <- data.frame(cbind(work, genders, ag, genre))
    chars <- data.frame(rbind(temp_chars, chars))
  }
  chars_20  <- data.frame(rbind(chars_20, chars))
  
  # chars_all <- data.frame(rbind(chars_all, chars))
}
```








```{r  include = FALSE}
# 1a: ADJUSTMENTS
# - - - - - - - - - 

# how many M / ? need to be replaced with Fs?

men_rep2  <- epi_func(chars_20$genders, 'M', 'F', '?', se_2m, sp_2m)
q_rep2  <- (epi_func(chars_20$genders, 'F', 'M', '?', se_2, sp_2))-abs(men_rep2)

# randomly replace M 

rep_m20 <- sample(which(chars_20$gender == 'M'), abs(men_rep2))
chars_20$adj_all <- replace(as.character(chars_20$gender), rep_m20, 'F')

# randomly replace ?

rep_q20 <- sample(which(chars_20$adj_all == '?'), q_rep2)
chars_20$adj_all <- replace(as.character(chars_20$adj_all), rep_q20, 'F')


# BOOTSTRAP 
# calculate the percentage of 'F' with bootstrapping 

boot_chars_20 <- vector()
boot_chars_all  <- vector()
for (j in 1:10000){  
  sample <- boot_func(chars_20$adj_all, chars_20)
  # sample  <- boot_func(chars_all$adj_all, chars_all)
  f_perc <- length(which(sample == 'F'))/length(sample)
  boot_chars_20[j] <- f_perc
  # boot_chars_all[j] <- f_perc
}

```

```{r  include = FALSE}

# 1b: REPORTED VALUES
# - - - - - - - - - - 

length(which(chars_20$genders == 'F')) # raw women
length(which(chars_20$adj_all == 'F')) # adjusted women 
length(which(chars_20$adj_all == 'F')) / nrow(chars_20) # % women 
mean(boot_chars_20) # bootstrap mean - top 20 
mean(boot_chars_all) # bootstrap mean - all 

# by genre and ag 

chars_func  <- function(x){
  results <- NULL
  for (i in 1:nlevels(x)){
    sub <- chars_20[which(as.character(x) == levels(x)[i]),]
    genre <- levels(x)[i]
    f <- vector()
    for (j in 1:10000){  
      sample <- boot_func(sub$adj_all, sub)
      f[j] <- length(which(sample == 'F'))
    }
    all  <- nrow(sub)
    temp <- data.frame(cbind('f'= mean(f), 'all'= all, genre))
    results <- data.frame(rbind(results,temp))
  }
  results$f_perc  <- num_func(results$f)/num_func(results$all)
  return(results)
}

# by genre and author gender 

boot_chars_genre <- chars_func(chars_20$genre)

# remove romance

chars_20xROM <- chars_20[which(chars_20$genre != 'ROM'),]
boot_chars_ag <- chars_func(chars_20xROM$ag)
boot_chars_ag_wROM <- chars_func(chars_20$ag)
```


```{r include=FALSE}

# 1d: SIGNIFICANCE TESTING
# - - - - - - - - - - - - - 

chisq_20  <- t(data.frame(cbind('women' = num_func(boot_chars_genre$f), 
                                'all'=num_func(boot_chars_genre$all))))
chisq.test(chisq_20)

chisq_20_ag  <- t(data.frame(cbind('women' = num_func(boot_chars_ag$f[1:2]),
                                   'all'=num_func(boot_chars_ag$all[1:2]))))
chisq.test(chisq_20_ag)
fisher.test(chisq_20_ag)
```


```{r results='asis'}

# 1e: TABLE TWO
# - - - - - - - - 

table_2  <- data.frame(cbind('genre' = as.character(boot_chars_genre$genre),
                             'percent_w' = num_func(boot_chars_genre$f_perc)*100))

table_2 <- table_2[order(num_func(table_2$percent_w)),]

table_2_tidy <- table_2 %>%
  gt() %>% 
  tab_header("Estimates for Prevalence of Women in Top-Twenty Characters in Novels of All Genres") %>% 
  cols_label(genre = "Genre",
             percent_w = "% Women Top-Twenty Characters")
print(table_2_tidy)

ggsave(filename = "table_2.jpg", plot = tableGrob(table_2), path = "./figures/", width = 7, height = 7)

```

```{r results='asis'}

# 1f: FIGURES THREE & FOUR
# - - - - - - - - - - - - -

# DATA 
# from ranks_sub, update the first and second character based on perspective calcs

ranks_sub[1,1:2]  <- c(623, 617)
ranks_sub[2,1:2]  <- c(470, 768)
ranks_sub_w[1,1:2]  <- c(321,194)
ranks_sub_w[2,1:2]  <- c(191,324)

# get correct percentages of women 

ranks_sub$perc_w  <- ranks_sub$w/ranks_sub$all
ranks_sub_w$perc_w  <- ranks_sub_w$w/ranks_sub_w$all

# BUILD 
# figure 3 
# uses only works with main character identified 

figure_3 <- ggplot(data=ranks_sub, aes(rank)) + 
  geom_line(aes(y = women, colour = "Women")) + 
  geom_line(aes(y = men, colour = "Men")) +
  scale_color_manual(values=c("darkorange2", "mediumpurple4")) +
  ggtitle("Men and Women by Rank Position (All Authors)")+
  labs(x="Rank of Character", y="") +
  labs(color="Character Gender") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, 
                                  # family = "Times New Roman", 
                                  face = "bold"),
        text = element_text(size = 11), 
                            # family = "Times New Roman"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) +
  geom_vline(xintercept = 20, linetype = "dashed") 

print(figure_3)

ggsave(filename = "figure_3.jpg", plot = figure_3, path = "./figures/", width = 7, height = 7)

```


``` {r results='asis'}

# figure 4 

figure_4 <- ggplot(data = ranks_sub_w, aes(rank)) + 
  geom_line(aes(y = women, colour = "Women")) + 
  geom_line(aes(y = men, colour = "Men")) +
  scale_color_manual(values=c("darkorange2", "mediumpurple4")) +
  ggtitle("Men and Women by Rank Position (Women Authors)")+
  labs(x="Rank of Character", y="") +
  labs(color="Character Gender") +
  theme_bw() +
  theme(plot.title = element_text(size = 13, 
                                  # family = "Times New Roman", 
                                  face = "bold"),
        text = element_text(size = 11), 
                            # family = "Times New Roman"),
        axis.title = element_text(face="bold"),
        axis.text.x=element_text(size = 11)) +
  geom_vline(xintercept = 20, linetype = "dashed") 

print(figure_4)

ggsave(filename = "figure_4.jpg", plot = figure_4, path = "./figures/", width = 7, height = 7)

```


